Plant-pollintor networks worldwide are composed by the same specific building

### 2021-10-16 Draft/Draft.Rmd
Conclusion and abstract

##Abstract

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consitently under- and over-represented worldwide, suggesting that the building blocks of plant-pollintor networks are not random. Interestingly, different functional groups are associated to specific motif position across both trophic levels. HERE I WOULD PUT AN EXAMPLE OF E.G. FAN POSITIONS, DEFINED BY, TEND TO BE OCCUPIED BY BEES AND WHATEVER TYPE OF PLANTS, RESPECTIVELY. This findings open the door to predict how novel interactions will be formed within a community context.

@@

##INTRODUCTION

@@

##METHODS

@@

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. For this, we used the trait dataset collated in Chapter 1 that comprised a total of 1506 species including the 503 species considered in this study (Supplementary material Chapter 1 **Table S1**). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30%; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. To feed the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) bees (Hymenoptera-Anthophila), (ii) non-bee Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Syrphidae-Diptera), (iv) non-syrphids-Diptera, (v) Lepidoptera and (vi) Coleoptera. However, a minor set of species belonged to other functional groups that were considered in analyses but not discussed further because of their low representation in the full set of networks (3.55% of the total interactions recorded). These functional groups were 'lizards', 'birds' and 'other insects'. This last group was formed by a mix of uncommon insect taxa on the full set of networks.

@@

Following previous work [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent. To test the effect of singletones, we ran all analysis to the same set of networks, but considering only interactions whose frequency was greater than one (64.98% of interactions). Results are qualitatively consistent and are not further discussed in the main text (**Figures S1 and S2**).

@@

Once we obtained $p_i^O$ and $p_i^E$, we used a simulation approach to determine whether the former is likely to come from the latter or not. This approach was preferred since the large number of possible motif combinations and the small probabilities for some of them advise against using an exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, denoted as $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. According to the usual interpretation of z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95%  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

@@

##RESULTS

@@

---
COMMENT
IB: Note that along the ms (abstract, discusion, etc...), I would discuss the group names ('selfing herbs') and not numbers (Group 1) otherwise we will lost the readers.
---

@@

---
THIS IS A COMMENT
Alfonso sobre la frase "These differences were more marked for floral visitors...": Cómo se infiere esto? Es viendo el dendrograma?

Jose: He modidficado el texto y creo que ahora está más claro
---

@@

The comparison of the functional group frequencies per motif position between empirical and simulated networks showed a `r over`% and `r under`% of over- (>75$^{th}$) and under- (<25th) represented functional groups in the different positions, respectively (**Figure 3**). Floral visitors showed a total of `r poll_over_75`% and `r poll_under_25`% of over- and under-represented functional groups in the different positions and plants `r plant_over_75`% and `r plant_under_25`% of over and under-represented groups, respectively. Notably, the differences across functional groups were more marked for floral visitors than for plant functional groups (the differences between min and max percentiles per position were generally two-three times larger for floral visitors). From most over- to under-represented floral visitors functional groups on the different motif positions (indicated with the dendrogram order in **Figure 3**), we found: bees, non-syrphids Diptera, syrphids, Coleoptera, non-bee Hymenoptera and Lepidoptera. Although plant functional groups showed less differences between them, there were also more represented functional groups than others, thus from most over- to under-represented groups on the different motif positions we found: self incompatible perennials with large flowers, small outcrossing perennials, tall plants with unisexual flowers, selfing herbs and short lived outcrossers with long zygomorphic flowers.

@@

The different observed frequencies on the different motif positions showed a hirarchical order of probabilities for floral visitors (**Figure 4**). That is, the functional group of 'bees' was the most abundant one on all motif positions followed by non-syrphid Diptera which was after bees, the second most frequent functional group on all positions. Then, the group of syrphids was the most abundant in all positions but two (positions 28 and 31 from motifs 11 and 12, respectively). After syrphids, Coleoptera was the most frequent group on the different positions followed by non-bee Hymenoptera and finally, Lepidoptera which was the less frequent group on all positions but two (positions 12 and 22 from motifs 5 and 9, respectively). Remarkably, plant functional groups were more variable on the different motif positions. However, there were also predominant functional groups found across the different motif positions. The three most frequent groups were 'tall plants with small unisexual flowers', 'self-incompatible perennials with large flowers' and 'small outcrossing perennials'. On the contrary, the groups of 'selfing herbs' and 'short lived outcrossers with long zygomorphic flowers' had the lowest probability to be present on the different motifs plant positions.

@@

The statistical comparison with z-scores between the observed and expected functional group combinations showed that we can not recover the observed combinations from its probability of ocurrence. In fact, most combinations are under- or over- representedm and follow a modified Gaussian distribution (Johnson's $S_U$) with 65% of under-represented group combinations, 20% of no statistical difference and 15% of under-represented ones (**Figure 5**). In addition, motifs with small node combinations (i.e., 2 and 3 nodes) only appeared as under-represented. Motifs with 4 and 5-node combinations appeared in the three statistical categories but 4 node motifs had a higher porportion in the under-represented category (10% in comparison with the 1% and the 3% of the no statiscal difference and over-represented categories, respectively). Finally, 5-node motifs were in similar prportions in the no statistical difference and over-represented categories (99% and 97%, respectively) but they appeared in lower porportion in the under-represented category (89%). When, we explore the identity on the different positions of the most probable motif combinations we find as expected that the most common functional groups rule these positions on the most probable motif combiantions (lower panel **Figure 5**).

@@

##DISCUSSION

@@

From a pollinator perpective, we found that generalist network motifs were consistently over-represented (**Figure 2**). For instance, when we compare motifs with equal structure but inverted, we found that floral visitors had always over-representation in the positions that implied the highest number of links. Despite this is in good agreement with the current view of the generalist tendency of pollinators [@waser1996; @olesen2002; @ollerton2017], we found that this over-representation of the generalist network motifs was largely driven by the functional group of bees (e.g., positions 42, 46, 16, 4, 14 and 35 in motifs 15, 17, 7, 2, 6 and 13, respectively; see **Figure 3** for positions and **Figure 4** for motifs) while the rest of the functional groups tended to be under-represented in these generalist motifs and over-represented in motif positions with specialist roles (e.g., positions 11, 21, 28 and 34 in motifs 5, 9, 11 and 13, respectively; see **Figure 3** for positions and **Figure 4** for motifs). This large effect of bees on the overall patterns is no surprise, as bees were the functional group with largest number of interaction records in the whole set of networks as also recorded for other broad scale plant-pollinator network studies [@carvalheiro2014; @ollerton2017]. Unlike floral visitors, plant functional groups showed little differences between their maximum and minimum percentile on the different motif positions in the comparison with simulated networks, and thus similar probability of appearing on them. However, the groups that involved higher levels of outcrossing showed a tendency to be more over-represented and probable on the different motif positions. For instance, the group of self-incompatible perennials with large flowers, one of the groups that is expected to have higher levels of outcrossing rate [@goodwillie2010], showed the highest over-representation of all plant functional groups (**Figure 3**); and the group of tall plants with unisexual flowers, the one that is expected to have the highest dependence of pollinators on reproduction [@rodger2021], was the most probable on the generalist motif for plant positions (i.e., plant nodes with 3 to 4 links; **Figure 4**).

The observed functional group probabilities on the different motif positions were insufficient to predict most of the observed network motifs (**Figure 5**). This implies that not just the representation of network motifs types but also the identity of the elements of their structure is influenced by ecological processes that determine under- and over-representation of the different motif combinations. In addition, we found that just 30% of the possible 53,250 motif combinations took place. Thus, similarly to the presence of forbidden links at the network level [@olesen2011; @jordano2016], we also found the existence of forbidden motifs that represented almost two thirds of the total combinations of indirect interactions. This concurs well with the general idea that plant-pollinator networks have a high porportion of non-realized direct interactions [@jordano2003; @bascompte2013], and as a result, also a high proportion of non-realized combinations of indirect interactions. Some of this non-observed interactions are probably due to under-sampling [@dorado2011; @jordano2016]. However, a large fraction of the possible interactions do not take place because of life history constraints such as species co-occurence, trait matching processes and phenological uncoupling [@olesen2011; @bartomeus2016; @jordano2016]. Despite the advances in the understanding on the ecological processes that influence species pairwise interactions [@bartomeus2016; @peralta2020], we still know little of how these processes determine competitive and facilitative interactions in mutualistic networks. Thus, as suggested by @simmons2020, exploring the proportion of these realized network motifs that is competitive or facilitative is a necessary next step to improve our understanding of the architecture of biodiversity [@bascompte2007].

Despite the obvious usefulness of funtional groups in summarizing species ecological roles (e.g., similar reproductive biology and behaviour of pollinators), delimiting species into functional groups is not often trivial. For instance, taxonomic diversity and trait availability limit our ability to properly represent the wide range of reproductive strategies [@dellinger2020] and different clustering analysis can lead to distinct outcomes (e.g., k-means and hierarchical clustering). Further, here we divided floral visitors on the taxonomic rank level because it represented adequately the pollinator species ecological roles but further groupings with trait data could help revealing greater relevance of trait-matching processes on indirect interactions. Moreover, we analysed network motifs focusing exclusively on presence-absence interactions but as suggested by @simmons2020, further work should account for quantitative measurements of interaction strength. Moreover, despite the wide geographical coverage of our plant-pollinator dataset, there are major gaps for some regions of the world with a clear bias towards Europe, America and islands which have long interested ecologist [@macarthur1963; @warren2015]. These biases are also found in global datasets of networks where the data is highly fragmented across space [@poisot2021] and more efforts are needed to capture global patterns of plant-pollinator networks.

In conclusion, by following the framework of @simmons2020 we have been able to explore competitive and facilitative interactions in plant-pollinator networks worldwide and found that indirect interactions are non-random and are more strongly related to neutral processes than any other ecological process. The association between network motifs and functional groups revealead that the combinations of functional groups is also strongly restricted by ecological processes

these groups rely on their ecological role (e.g., generalization versus specialization) and relies also

Further, the different network motifs are associated with different fucntional groups which translates into different ecological roles of these functional groups at macroecological scale.

with specific network motif positions associated to different species ecological roles. This work set the stage to help predicting novel interactions beyond pairwise interactions, to explore the overall importance

Further, bees were the only generalist functional group in comparison with the rest floral visitor groups that appeared more frequently in specialist network motifs. We also found that the the node position was partly associated to the ecological role of plants but without large differences between groups. Our work has explored for the first time worldwide plant-pollinator patterns of indirect interactions and opens the door for predicting novel interactions beyond pairwise interactions.

Final one: 
5) Future direction and conclusions (also short). This opens the door to help predicting novel interactions beyond pairwise interactions. We not only know now how the building blocks are structured but also that who conform this building blocks is not random.

Idea: explore if over-represented and under-represented groups have trait-matching or not... As suggested in the literature the more stable (and probably more frequent I guess will have well matched traits...)

##REFERENCES

>>>>

# Abstract

Ecological processes leave distinct structural imprints on species interactions shaping the topology of mutualistic networks. Detecting those relationships is not trivial since they go beyond pair-wise interactions, but may get blurred when considering full network descriptors. However, recent work has shown the network meso-scale can capture this important information. The meso-scale describes network subgraphs representing patterns of interactions between a small number of species (i.e. motifs) and those constitute the building blocks of the whole network. Here, we have compiled 60 networks from 18 different studies and show that some motifs are consistently under- and over-represented worldwide, suggesting that the building blocks of plant-pollinator networks are not random and are most influenced by species relative abundances. Interestingly, network motif combinations are restricted to almost two thirds of all the potential indirect interactions and we found that distinct positions that imply different ecological roles (e.g., specialization versus generalization) are occupied by different functional groups on both trophic levels. For instance, bees were consistently on generalist motif positions while the rest of functional groups tended to appear on specialist ones. Although for plants these differences were less marked, we found that the species with higher levels of outcrossing had the most number of indirect interactions. Our result highlight the non-random structure of indirect interactions in plant-pollinator networks and opens the door to predict novel interactions beyond pairwise interactions within a community context.

@@

# INTRODUCTION

@@

# METHODS

@@

First, plant species were grouped through hierarchical clustering into the optimal number of functional groups that summarized the main plant reproductive strategies. For this, we used the trait dataset collated in Chapter 1 that comprised a total of 1506 species including the 503 species considered in this study (Supplementary material Chapter 1 **Table S1**). This dataset consisted on 8 floral, 4 reproductive and 3 vegetative traits excluding traits with high percentage of missing values (over 30$\%$; **Table S2**). We opted to calculate the plant functional groups on this larger set of species because of the higher accuracy when delimiting functional groups with that many variables and species [@dolnicar2014]. To feed the clustering analysis, we calculated the distance between the different qualitative and scaled quantitative variables with Gower distance [@gower1971]. For this, we used the function _gowdis_ with method _ward.D2_ from the package _FD_ version _1.0-12_ [@laliberte2014]. Finally, we conducted hierarchical clustering with the function _hclust_ from the _R_ stats package version _4.0.5_ and calculated the optimal number of clusters with the function _kgs_ from the package _maptree_ version _1.4-7_ [@white2009].

Second, floral visitors were grouped into functional groups based on taxonomic rank as done similarly in past plant-pollinator studies [@fenster2004; @ollerton2009]. We opted to divide floral visitors on the taxonomic rank level and not with functional traits because the main orders of floral visitors differed in form and behaviour and had lower superior taxonomical complexity (i.e., floral visitors had 6 orders versus plants that had 38). Thus, this allowed us to group floral visitors into functional groups that represented adequately the main life strategies of the possible pollinators: (i) bees (Hymenoptera-Anthophila), (ii) non-bee Hymenoptera (Hymenoptera-non-Anthophila), (iii) syrphids (Syrphidae-Diptera), (iv) non-syrphids-Diptera, (v) Lepidoptera and (vi) Coleoptera. However, a minor set of species belonged to other functional groups that were considered in analyses but not discussed further because of their low representation in the full set of networks (3.55$\%$ of the total interactions recorded). These functional groups were 'lizards', 'birds' and 'other insects'. This last group was formed by a mix of uncommon insect taxa on the full set of networks.

@@

Following previous work [@simmons2019; @simmons2020], we broke down the plant-pollinator networks into their constituent motifs. Prior to analyses, we turned the quantitative networks into qualitative (or binary) ones, where interactions are present or absent. To test the effect of singletones, we ran all analysis to the same set of networks, but considering only interactions whose frequency was greater than one (64.98$\%$ of interactions). Results are qualitatively consistent and are not further discussed in the main text (**Figures S1 and S2**).

@@

Once we obtained $p_i^O$ and $p_i^E$, we used a simulation approach to determine whether the former is likely to come from the latter or not. This approach was preferred since the large number of possible motif combinations and the small probabilities for some of them advise against using an exact test of goodness-of-fit or a Chi-square one. Specifically, we created 1,000 random samples with repetition of possible motif combinations, where each sample contained 10 million elements and, for each combination, the probability of being selected was equal to its expected probability. From those random samples, we extracted the mean and the standard deviation of the expected probability of $i$, denoted as $E\left [ p_i^E \right ]$ and $\sigma\left [ p_i^E \right ]$, respectively, and calculated the z-scores of $p_i^O$ as $z_i^O = \left ( p_i^O - E\left [ p_i^E \right ] \right )/\sigma \left [ p_i^E \right ]$, for those motif combinations with $p_i^O > 0$. According to the usual interpretation of z-scores, combinations with $z_i^O > 1.96$ are over-represented, whereas those with $z_i^O < -1.96$ are under-represented, at the 95$\%$  confidence level. Notice that we focused on combinations with $p_i^O > 10^{-7}$ because, beyond the interest that can arouse the combinations that appear at most once, we do not have enough numerical resolution to accurately detect whether or not those combinations are under-represented (due to the limited size of our random samples).

@@

#RESULTS

@@

@@

@@

The comparison of the functional group frequencies per motif position between empirical and simulated networks showed a `r over`$\%$ and `r under`$\%$ of over- (>75$^{th}$) and under- (<25th) represented functional groups in the different positions, respectively (**Figure 3**). Floral visitors showed a total of `r poll_over_75`$\%$ and `r poll_under_25`$\%$ of over- and under-represented functional groups in the different positions and plants `r plant_over_75`$\%$ and `r plant_under_25`$\%$ of over and under-represented groups, respectively. Notably, the differences across functional groups were more marked for floral visitors than for plant functional groups (the differences between min and max percentiles per position were generally two-three times larger for floral visitors). From most over- to under-represented floral visitors functional groups on the different motif positions (indicated with the dendrogram order in **Figure 3**), we found: bees, non-syrphids Diptera, syrphids, Coleoptera, non-bee Hymenoptera and Lepidoptera. Although plant functional groups showed less differences between them, there were also more represented functional groups than others, thus from most over- to under-represented groups on the different motif positions we found: self incompatible perennials with large flowers, small outcrossing perennials, tall plants with unisexual flowers, selfing herbs and short lived outcrossers with long zygomorphic flowers.

@@

**Over- and under-represented functional group combinations of motifs**

The different observed frequencies on the different motif positions showed a hierarchical order of probabilities for floral visitors (**Figure 4**). That is, the functional group of 'bees' was the most abundant one on all motif positions followed by non-syrphid Diptera which was after bees, the second most frequent functional group on all positions. Then, the group of syrphids was the most abundant in all positions but two (positions 28 and 31 from motifs 11 and 12, respectively). After syrphids, Coleoptera was the most frequent group on the different positions followed by non-bee Hymenoptera and finally, Lepidoptera which was the less frequent group on all positions but two (positions 12 and 22 from motifs 5 and 9, respectively). Remarkably, plant functional groups were more variable on the different motif positions. However, there were also predominant functional groups found across the different motif positions. The three most frequent groups were 'tall plants with small unisexual flowers', 'self-incompatible perennials with large flowers' and 'small outcrossing perennials'. On the contrary, the groups of 'selfing herbs' and 'short lived outcrossers with long zygomorphic flowers' had the lowest probability to be present on the different motifs plant positions.

@@

The statistical comparison with z-scores between the observed and expected functional group combinations showed that we can not recover the observed combinations from its probability of occurrence. In fact, most combinations are under- or over- represented and follow a modified Gaussian distribution (Johnson's $S_U$) with 65$\%$ of under-represented group combinations, 20$\%$ of no statistical difference and 15$\%$ of under-represented ones (**Figure 5**). In addition, motifs with small node combinations (i.e., 2 and 3 nodes) only appeared as under-represented. Motifs with 4 and 5-node combinations appeared in the three statistical categories but 4 node motifs had a higher proportion in the under-represented category (10% in comparison with the 1$\%$ and the 3$\%$ of the no statistical difference and over-represented categories, respectively). Finally, 5-node motifs were in similar proportions in the no statistical difference and over-represented categories (99$\%$ and 97$\%$, respectively) but they appeared in lower proportion in the under-represented category (89$\%$). When, we explore the identity on the different positions of the most probable motif combinations we find as expected that the most common functional groups rule these positions on the most probable motif combinations (lower panel **Figure 5**).

@@

# DISCUSSION

@@

From a pollinator perspective, we found that generalist network motifs were consistently over-represented (**Figure 2**). For instance, when we compare motifs with equal structure but inverted, we found that floral visitors had always over-representation in the positions that implied the highest number of links. Despite this is in good agreement with the current view of the generalist tendency of pollinators [@waser1996; @olesen2002; @ollerton2017], we found that this over-representation of the generalist network motifs was largely driven by the functional group of bees (e.g., positions 42, 46, 16, 4, 14 and 35 in motifs 15, 17, 7, 2, 6 and 13, respectively; see **Figure 3** for positions and **Figure 4** for motifs) while the rest of the functional groups tended to be under-represented in these generalist motifs and over-represented in motif positions with specialist roles (e.g., positions 11, 21, 28 and 34 in motifs 5, 9, 11 and 13, respectively; see **Figure 3** for positions and **Figure 4** for motifs). This large effect of bees on the overall patterns is no surprise, as bees were the functional group with largest number of interaction records in the whole set of networks as also recorded for other broad scale plant-pollinator network studies [@carvalheiro2014; @ollerton2017]. Unlike floral visitors, plant functional groups showed little differences between their maximum and minimum percentile on the different motif positions in the comparison with simulated networks, and thus similar probability of appearing on them. However, the groups that involved higher levels of outcrossing showed a tendency to be more over-represented and probable on the different motif positions. For instance, the group of self-incompatible perennials with large flowers, one of the groups that is expected to have higher levels of outcrossing rate [@goodwillie2010], showed the highest over-representation of all plant functional groups (**Figure 3**); and the group of tall plants with unisexual flowers, the one that is expected to have the highest dependence of pollinators on reproduction [@rodger2021], was the most probable on the generalist motif for plant positions (i.e., plant nodes with 3 to 4 links; **Figure 4**).

The observed functional group probabilities on the different motif positions were insufficient to predict most of the observed network motifs (**Figure 5**). This implies that not just the representation of network motifs types but also the identity of the elements of their structure is influenced by ecological processes that determine under- and over-representation of the different motif combinations. In addition, we found that just 30$\%$ of the possible 53,250 motif combinations took place. Thus, similarly to the presence of forbidden links at the network level [@olesen2011; @jordano2016], we also found the existence of forbidden motifs that represented almost two thirds of the total combinations of indirect interactions. This concurs well with the general idea that plant-pollinator networks have a high proportion of non-realized direct interactions [@jordano2003; @bascompte2013], and as a result, also a high proportion of non-realized combinations of indirect interactions. Some of this non-observed interactions are probably due to under-sampling [@dorado2011; @jordano2016]. However, a large fraction of the possible interactions do not take place because of life history constraints such as species co-occurrence, trait matching processes and phenological uncoupling [@olesen2011; @bartomeus2016; @jordano2016]. Despite the advances in the understanding on the ecological processes that influence species pairwise interactions [@bartomeus2016; @peralta2020], we still know little of how these processes determine competitive and facilitative interactions in mutualistic networks. Thus, as suggested by @simmons2020, exploring the proportion of these realized network motifs that is competitive or facilitative is a necessary next step to improve our understanding of the architecture of biodiversity [@bascompte2007].

Despite the obvious usefulness of functional groups in summarizing species ecological roles (e.g., similar reproductive biology and behaviour of pollinators), delimiting species into functional groups is not often trivial. For instance, taxonomic diversity and trait availability limit our ability to properly represent the wide range of reproductive strategies [@dellinger2020] and different clustering analysis can lead to distinct outcomes (e.g., k-means and hierarchical clustering). Further, here we divided floral visitors on the taxonomic rank level because it represented adequately the pollinator species ecological roles without the need of the tremendous effort of compiling trait data for over 1000 species of floral visitors. However, further groupings with traits could help revealing greater relevance of trait-matching processes on indirect interactions. We believe that this could be investigated by exploring which proportion of over- and under-represented motifs present high and low trait-matching, respectively. Moreover, we analysed network motifs focusing exclusively on presence-absence interactions which assume that all interactions are equally important but as suggested by @simmons2020, further work should account for quantitative measurements of the strength of plant-pollinator interactions. Moreover, despite the wide geographical coverage of our plant-pollinator dataset, there are major gaps for some regions of the world with a clear bias towards Europe, America and islands which have long been interested ecologist [@macarthur1963; @warren2015]. These biases are also found in global datasets of networks where the data is highly fragmented across space [@poisot2021] and more efforts are needed to capture global patterns of plant-pollinator interactions.

In conclusion, by following the framework of @simmons2020 we have been able to explore competitive and facilitative interactions in plant-pollinator networks worldwide and found that indirect interactions are non-random and are more strongly related to neutral processes than any other ecological process. Further, we found that motif combinations of functional groups are strongly constrained and that the distinct motif positions are associated with different functional groups with different ecological roles (e.g., generalization versus specialization). Although in this study we do not predict indirect interactions, our work is a first step to elucidate novel interactions beyond pairwise interactions.

# REFERENCES
